Automated Lensing Learner - I: An automated Strong Lensing Identification Pipeline

### 2017-03-21 ms.tex
add nan's text for the simulation section

%-------------------------------------------------%
\section{Methodology}
\label{sec:methods}
%-------------------------------------------------%
\subsection{Mock Observations}
%-------------------------------------------------%
\subsection{Mock Images}
%-------------------------------------------------%
% Summary of HST and LSST images
We use mock {\em Hubble Space Telescope}\footnote{\todo{hst...org}}
(HST) and {\em Large Synoptic Sky
  Telescope}\footnote{\todo{lsst....org}} (LSST) images generated with
PICS (Pipeline for Images of Cosmological Strong lensing)
\citep{li_etal16} and LensPOP \citep{collett_15} to train and test our
model.  Along the line of sight galaxies and foreground stars.  Half
of the images used are strong lensing systems.  Lensed galaxy images
are ray-traced images of actual galaxies extracted from deep Hubble
Space Telescope observations.  The final mock images are ``observed''
with a realistic point spread function corresponding to modeled
detector artifacts for bright stars.  We refer to \cite{li_etal16} for
further details, but outline the process here.

We generate 10,000 lens containing mock observations and 10,000
non-lens containing mock observations to run our parameter search, and
keep a hold-out test set of 1,000 lens containing mock observations
and 1,000 non-lens containing mock observations on which we evaluate
the final trained model.

Lens containing mock observations include the lens galaxy, lensed
images of the source galaxy, and galaxies along the line of sight.
Non-lens containing mock observations include all but a lensed source
image.  We convolve each of the $2\times 10,000$ train/test images and
$2\times 1,000$ hold-out images to produce three separate sets of mock
``observations'' with equal numbers of lens and non-lens containing
systems: (1) HST-like observations, (2) LSST-like observations over
the span of one year, and (3) LSST-like observations over the span of
ten years.  We respectively label these observations as HST, LSST1,
and LSST10.
%-------------------------------------------------%
\subsubsection{Modeling Images of the Source Galaxies}
%-------------------------------------------------%
To produce images of the simulated lensed source galaxies, we first
set mass models of the lens galaxy to be the Singular Isothermal
Ellipsoid (SIE):
%-------------------------------------------------%
\begin{equation}\label{eqn:sie}
\kappa = \frac{\theta_{\rm E}}{2}\frac{1}{\sqrt{x_1^2/q+x_2^2 q}},
\end{equation}
%-------------------------------------------------%
where $\theta_{\rm E}$ is the Einstein radius and $q$ is the axis
\wording{ratio between semi-major and semi-minor axes $x_1$ and
  $x_2$}. We can calculate the Einstein radius from the redshift of
the lens, the redshift of the source, and the velocity dispersion of
the lens galaxy.
%-------------------------------------------------%
\begin{equation}\label{eqn:einsteinradius}
\theta_{\rm E} = 4\pi\left(\frac{\sigma_v}{c}\right)\frac{D_{ls}}{D_{s}}. 
\end{equation}
%-------------------------------------------------%
Here, $c$ is the speed of light, $\sigma_v$ is the velocity dispersion
of the lens galaxy, and $D_{ls}$ and $D_{s}$ are the respective
angular diameter distance from the source plane to the lens plane and
from the source plane to the observer.

To rotate the lenses with a random orientation angles, we adopt the
transformation below for orientation angle, $\phi$:
%-------------------------------------------------%
\begin{eqnarray}
\label{eqn:rotation}
\begin{bmatrix}

    x_{1}\prime      \\
    x_{2}\prime 
\end{bmatrix}
= 
\begin{bmatrix}
    \cos \phi  &  -\sin \phi      \\
    \sin \phi  &  ~\cos \phi      
\end{bmatrix} 
\begin{bmatrix}
    x_1      \\
    x_2      
\end{bmatrix} .
\end{eqnarray}
%-------------------------------------------------%
From Equations~\ref{eqn:sie}~-~\ref{eqn:rotation}, there are five
independent parameters that uniquely define the lensing system:
$\sigma_v$, ellipticity $q$, orientation angle $\phi$, redshift of the
lens $z_l$, and redshift of the source $z_s$.

% Parameters of lensing system
We choose $\sigma_v$, $q$, and $\phi$ randomly from typical ranges of
a galaxy, where $\sigma_v = [200, 320] km/s$, $q = [0.5, 1.0]$, and
$\phi = [0, 360]$.  The redshift of the lens galaxy is drawn from the
distribution \todo{found in \citet{???}}.  We fix the redshift of the
source to $z_s = 1.0$. We then extract images of 33 bright galaxies
from the CANDELS survey as source galaxies.  We place the source
galaxy images in projected positions close to the caustic of the
lensing system and fix the projected position of the lens galaxy to
the center of the field of view of each image.  Once we have
determined all parameters of the lensing system, we perform
ray-tracing simulations to produce the final lensed image.

%-------------------------------------------------%
\subsubsection{Modeling Lens Galaxies}
%-------------------------------------------------%
We model the light distributions of the lens galaxies with an
elliptical Sersic profile,
%-------------------------------------------------%
\begin{equation}\label{eqn:sersic}
I(R) = I_{\rm eff}~{\rm exp} \left\lbrace -b_{n} \left[ \left(
  \frac{R}{R_{\rm eff}}\right)^{1/n} - 1 \right ] \right\rbrace,
\end{equation}
%-------------------------------------------------%
where, $R = \sqrt[]{x_1^2 /q+x_2^2 q }$, $R_{\rm eff}$ is the
effective radius in units of arcseconds, $I_{\rm eff}$ is the
intensity at the effective radius, $n$ is the index of the Sersic
profile, and $q$ is the ellipticity of the lens galaxy.  We perform a
similar transformation as Equation~\ref{eqn:rotation} to orient the
sources.  We assume that the distribution of light is following that
of mass, so ellipticity and orientation are the same as the SIE model.

@@

%-------------------------------------------------%
\subsubsection{Producing Mock Images}
%-------------------------------------------------%

The final step is to perform a mock observations.  For HST-like
observations, galaxies along the line of sight already include noise
and a point spread function (PSF).  We therefore stack and magnitude
calibrate all components to produce the final HST-like images, which
are $300\times300~{\rm pix}^2$ with $0.03$~arcsec per pixel.

To produce LSST-like images, we use {\em LensPop} to serially perform
re-binning, convolve with the corresponding PSF, and add
noise\citep{collett_15}.  We created one year LSST-like (LSST1) images
and stacked ten-year LSST-like images (LSST10) to investigate the
performance and usability of our pipeline in future LSST
data. LSST-like images are $45\times45~{\rm pix}^2$ with $0.2$~arcsec
per pixel.

@@

>>>>

%-------------------------------------------------%
\section{Methodology}
\label{sec:methods}
We use mock {\em Hubble Space Telescope}\footnote{\todo{hst.org}} (HST) and {\em Large
Synoptic Sky Telescope}\footnote{\todo{lsst.org}}
(LSST) images generated with PICS (Pipeline for Images of Cosmological Strong
lensing) \citep{li_etal16} and LensPop\citep{collett_etal15} for training and testing.

There are 10,000 lens containing mock observations and 10,000
non-lens containing mock observations for running our parameter search. 
We keep a hold-out test set of 1,000 lens containing mock observations
and 1,000 non-lens containing mock observations on which we evaluate
the final trained model.

Mock lensing systems contain mock observations include the lens galaxy, lensed
images of the source galaxy, and galaxies along the line of sight.
Mock Non-lensing systems contain mock observations include all but a lensed source
image.  We convolve each of the $2\times 10,000$ train/test images and
$2\times 1,000$ hold-out images to produce three separate sets of mock
``observations'' with equal numbers of lens and non-lens containing
systems: (1) HST-like observations, (2) LSST-like observations over
the span of one year, and (3) LSST-like observations over the span of
ten years.  We respectively label these observations as HST, LSST1,
and LSST10.
%-------------------------------------------------%
\subsubsection{Modeling the Mass Distribution of Lens Galaxies}
To produce simulated lensed galaxies, we first set mass models of the lenses to
be the Singular Isothermal Ellipsoid (SIE). This model is not only analytically
tractable, it has also been found to be consistent with models of individual
lenses and lens statistics on the length scales relevant for strong lensing
(e.g. Koopmans et al. 2006\footnote{Koopmans L. V. E., Treu T., Bolton A. S.,
Burles S., Moustakas L. A., 2006, The Astrophysical Journal, 649, 599}; Gavazzi
et al. 2007\footnote{Gavazzi R., Treu T., Rhodes J. D., Koopmans L. V. E.,
Bolton A. S., Burles S., Massey R. J., Moustakas L. A., 2007, The Astrophysical
Journal, 667, 176}; Dye et al. 2008\footnote{Dye S., Evans N. W., Belokurov V.,
Warren S. J., Hewett P., 2008, Monthly Notices of the Royal Astronomical
Society, 388, 384}; Li et al.
2009\footnote{http://iopscience.iop.org/article/10.1088/1674-4527/9/11/001/meta}).
The normalized convergence map of SIE model is defined as :

\begin{equation}
\kappa = \frac{\theta_{\rm E}}{2}\frac{1}{\sqrt{x_1^2/q+x_2^2 q}}.
\end{equation}

Where, $\theta_{\rm E}$ is the Einstein Radius, $q$ is axis
ratio. Einstein radius can be calculated according to the redshift of
the lens, the redshift of the source, and the velocity dispersion of
the lens galaxy.
\begin{equation}
\theta_{\rm E} = 4\pi\left(\frac{\sigma_v}{c}\right)\frac{D_{ls}}{D_{s}}, 
\end{equation}
where, $c$ is the speed of light, $\sigma_v$ is the velocity
dispersion of the lens galaxy, $D_{ls}$ and $D_{s}$ are the angular
diameter distance from the source plane to the lens plane and from the
source plane to the observer separately.

To rotate the lenses with a random orientation angles, we adopt the
transformation below:
\begin{eqnarray}
\label{eq:rotation}
\begin{bmatrix}
    x_{1}\prime      \\
    x_{2}\prime 
\end{bmatrix}
= 
\begin{bmatrix}
    \cos \phi  & -\sin \phi      \\
    \sin \phi  & ~\cos \phi      
\end{bmatrix} 
\begin{bmatrix}
    x_1      \\
    x_2      
\end{bmatrix}.
\end{eqnarray}
$\phi$ is the orientation angle.  According to above equations, we need
$\sigma_v$, $q$, $\phi$, redshift of the lens $z_l$, and redshift of the source
$z_s$ to derivative the lensing maps completely. Velocity dispersion,
ellipticity, and orientation are randomly chosen from typical ranges of a
galaxy, where $\sigma_v = [150, 350] km/s$, $q = [0.5, 1.0]$, and $\phi = [0,
360]$ degrees. The lens redshifts are obtained by matching the velocity
dispersion in our simulations to the catalog of elliptical galaxies in the
COSMOS survey from Zahid et al. (2015)\footnote{Zahid H. J., Damjanov I.,
Geller M. J., Chilingarian I., 2015, The Astrophysical Journal, 806, 122}. The
resulting redshift range of our lenses is zl ∈ [0.2, 0.7].

%-------------------------------------------------%
\subsubsection{Modeling Images of the Lens Galaxies}
%-------------------------------------------------%
We model the light distributions of the lens galaxies with an
elliptical Sersic profile,
%-------------------------------------------------%
\begin{equation}\label{eqn:sersic}
I(R) = I_{\rm eff}~{\rm exp} \left\lbrace -b_{n} \left[ \left(
  \frac{R}{R_{\rm eff}}\right)^{1/n} - 1 \right ] \right\rbrace,
\end{equation}
%-------------------------------------------------%
where, $R = \sqrt[]{x_1^2 /q+x_2^2 q }$, $R_{\rm eff}$ is the
effective radius in units of arcseconds, $I_{\rm eff}$ is the
intensity at the effective radius, $n$ is the index of the Sersic
profile, and $q$ is the ellipticity of the lens galaxy.  We perform a
similar transformation as Equation~\ref{eqn:rotation} to orient the
sources.  We assume that the distribution of light is following that
of mass, so ellipticity and orientation are the same as the SIE model.

@@

The light distributions of the lens galaxies  are modeled as the elliptical
Sersic profile, \begin{equation} I(R) = I_{\rm eff}~{\rm exp} \left\lbrace
-b_{n} \left[ \left( \frac{R}{R_{\rm eff}}\right)^{1/n} - 1 \right ]
\right\rbrace \end{equation} where $R = \sqrt[]{x_1^2 /q+x_2^2 q }$, $R_{\rm
eff}$ is the effective radius in units of arcsecond, $I_{\rm eff}$ is the
intensity at the effective radius, $n$ is the index of the Sersic profile, $q$
is the ellipticity, and We perform the similar transformation as Eq.
\ref{eq:rotation} to orient the sources.  We assume that the distribution of
light is following that of mass, so ellipticity and orientation are the same as
the SIE model. Matching the velocity dispersion with the COSMOS morphological
catalog \footnote{\url{https://arxiv.org/pdf/1501.04977v2.pdf}}, we assign
effective radius, effective luminosity, and index to the light profile.  We
also assume the light center is on top of the mass center, the noiseless images
of lenses can be drawn. The galaxies on the line of sight are constructed by
cutting light-cones from the Hubble Ultra Deep Field.  Stacking all these
images together and collaborating the magnitude of each component, we create
primarily simulated lensed images.

\subsubsection{Modeling Images of the Source Galaxies}

The background sources are a set of detailed images of low-redshift bright
galaxies (z ∼ 0.45) extracted from the mosaics produced by the CANDELS team
(Grogin et al. 2011\footnote{Grogin N. A., et al., 2011, The Astrophysical
Journal Supplement Series December The Astrophysical Journal Supplement Series,
197, 3535}; Koekemoer et al. 2011\footnote{Koekemoer A. M., et al., 2011, The
Astrophysical Journal Supplement Series December The Astrophysical Journal
Supplement Series, 197}) and selected from the CANDELS UDS catalog Galametz et
al. (2013)\footnote{Galametz A., et al., 2013, The Astrophysical Journal
Supplement Series, 206, 10}. These galaxies are rescaled and placed at a fixed
redshift $z_s = 2.0$, near the caustics of the lensing system, and the
projected position of the lens is fixed in the center of the field of view of
each image. Therefore, once we have the lens, source, and the parameters of the
image simulation, i.e., $300 \times 300 {\rm pixels}^2$ and 0.03 arcsec per
pixel per side, settled, the lensed images can be produced by performing
ray-tracing simulations.

%-------------------------------------------------%
\subsubsection{Mock Observing}
%-------------------------------------------------%

The final part is mock observing. For HST-like
observations, galaxies along the line of sight already include noise
and a point spread function (PSF).  We therefore stack and magnitude
calibrate all components to produce the final HST-like images, which
are $300\times300~{\rm pix}^2$ with $0.03$~arcsec per pixel.

To produce LSST-like images from the ideal images produced in the previous steps,
we use the LensPop software Collett (2015)\footnote{Collett T. E., 2015, The
Astrophysical Journal, 811, 20} to perform mock observing. Images are resampled
to match the detector pixel scale and convolved with a circularly symmetric
Gaussian Point Spread Function discretised at the same pixel scale. A noisy
realisation of this image is generated assuming a Poisson model based on the
sky plus signal, and an additional Gaussian read-out noise. Parameters for
these simulations follow Collett (2015)\footnote{Collett T. E., 2015, The
Astrophysical Journal, 811, 20} and are based on the LSST observation simulator
(Connolly et al. 2010)\footnote{Connolly A. J., et al., 2010, SPIE Astronomical
Telescopes + Instrumentation, pp 77381O–77381O–10}. To account for seeing and
sky-brightness variations over the course of the survey, each simulated
exposure is drawn from a stochastic distribution of these parameters. We then
consider two different strategies to use these exposures. For each field, we
build one single-epoch image by keeping only the best seeing exposure and
another ``worst-case''' stack image by co-adding all individual exposures after
degrading them all in resolution to match the one with the worst seeing. These
two sets of images will allow us to investigate the trade-off between
signal-to-noise and resolution for our automated lens search. Figure 2.1
illustrates a sample mock lensed and unlensed images from each telescope. From
left to right, the top panel images respectively correspond to a mock HST
lensing and non-lensing system. The mock HST images are 300 × 300 in size, and
the pixel size is 0.03 arcsecond. For the HST-like data-set, many giant arcs
are visually obvious. The middle panel images respectively correspond to a mock
single-epoch LSST lensing and non-lensing systems. The bottom panel images
respectively correspond to a mock stacked LSST lensing and non-lensing systems.
These images are 45 × 45; the pixel size is 0.18 arcsecond. The ground based
noise, PSF, and limited resolution of the LSST-like data-set make visual giant
arc identification difficult. But when in the stacked LSST images, the quality
is better, because more data improves the signal to noise ratio better.

@@

In the training sets, we also generate unlensed images, the procedure is
similar to mock lensed images but getting rid of the part of ray-tracing. 10000
lensed images and 10000 unlensed images are merged to construct one training
set. Furthermore, for investigating the influence of BCG galaxies on the
performance of our lens identification pipeline in LSST data, we build another
training sets of the images without BCG galaxies.
